Imputation of incomplete multilevel data

### 2024-02-29 manuscript.qmd
add incomplete data to manuscript

>>>>

The most general analytic model to be applied to imputed data (Hox et al, 2018).

\begin{align}
\text{popularity}_{ij} = &\gamma_{00}+ \gamma_{10} \text{gender}_{ij} + \gamma_{20} \text{extraversion}_{ij} + \\
&\gamma_{01} \text{experience}_j + \gamma_{21} \text{extraversion}_{ij} \times \text{experience}_j + \\
&u_{2j} \text{extraversion}_{ij} + u_{0j}+ e_{ij} 
\end{align}

<!-- u_{1j} \text{gender}_{ij} +  -->

Setup `R` environment
```{r}
library(mice)     # for imputation
library(ggmice)   # for visualization
library(ggplot2)  # for visualization
library(dplyr)    # for data manipulation
library(lme4)     # for multilevel modeling
```

Load the data
```{r}
load("data/popular_MAR.RData")
```

Inspect the missingness
```{r}
plot_pattern(popular_MAR)
```

Fix cluster-level missingness determinately
```{r}
meth <- make.method(popular_MAR)
meth["experience_j"] <- "2lonly.mean"
meth[meth != "2lonly.mean"] <- ""
pred <- make.predictorMatrix(popular_MAR)
pred[, "cluster_id"] <- -2
imp <- mice(
  popular_MAR, 
  m = 1, 
  method = meth, 
  predictorMatrix = pred,
  maxit = 1)
plot_pattern(complete(imp))
```

Does not work if there are inconsistencies within clusters! But we can do this manually.

```{r}
popular_MAR |>
  group_by(cluster_id) |>
  reframe(teacher_experience = unique(experience_j))
```

This shows us that in the first cluster teacher experience of 4 and 24 is recorded.

```{r}
popular_MAR |>
  filter(cluster_id == 1) |>
  ggplot(aes(experience_j)) + geom_histogram()
```

For most pupils, teacher experience is 24. We can quite safely assume that the teacher expeience of 4 is a data entry error. We can fix this by over-imputing the teacher experience of 4 to 24, and deductively imputing the missing teacher experience from the same cluster as 24.
